\documentclass[]{article}
\usepackage{amsmath}		% For generic math symbols
\usepackage{amssymb}		% For mathbb
\usepackage{enumerate}		% For lists indexed by letters
\usepackage{bm}				% For bold letters
\usepackage{enumitem}		% So we can resume counting problem numbers after
% interrupting with text
\usepackage{hyperref}		% For clickable URL links
\usepackage{url}			% So file names won't create hboxes
\usepackage[margin=1in]{geometry}	% Make margins wider


\setlength{\parindent}{0pt}	% Turns off indentation


% Set some useful commands
\newcommand{\half}{\frac{1}{2}}			% 1/2
\newcommand{\R}{\mathbb{R}}				% Reals symbol
\newcommand{\bbm}{\begin{bmatrix}}		% Begin bmatrix environment
\newcommand{\ebm}{\end{bmatrix}}		% End bmatrix environment
\newcommand{\x}{\bm{x}}					% Bold (vector) x
\newcommand{\y}{\bm{y}}					% Bold (vector) y
\newcommand{\A}{\bm{A}}					% Bold (matrix) A
\newcommand{\vspan}{\mathrm{span}}		% To use the word span in math mode
\newcommand{\vrange}{\mathrm{range}}		% To use the word span in math mode
\newcommand{\vdim}{\mathrm{dim}}		% To use the word span in math mode
\newcommand{\vnull}{\mathrm{null}}		% To use the word span in math mode
\newcommand{\vrank}{\mathrm{rank}}		% To use the word span in math mode
\newcommand{\la}{\langle}				% Left angled bracket <
\newcommand{\ra}{\rangle}				% Right angled bracket >

% Place this command after each problem, before solution (examples below)
\newcommand{\solution}{\vskip 0.5cm \textbf{\large Solution:} \\}


\title{AMATH 352: Problem Set 6}
\author{Dave Moore, dmmfix@uw.edu}

\begin{document}

\maketitle
    {\Large \textbf{Due: Friday February 24, 2017}} \\

    \section*{Norms and Inner products:}
    \begin{enumerate}[resume]
	\item Let $\bm{W}$ be an invertible matrix. Show that the map
	  \[
	  \|\x\|_{\bm{W}}=\|\bm{Wx}\|_2
	  \]
	  is a norm on $\R^m$. Is it still a norm if $\bm{W}$ is singular? Why or why not?

	  \solution Since $\bm{W}$ is invertible, its range is all of
      $\R^m$, which means that we can simply observe that
      $\|\cdot\|_2$ is a norm on $\R^m$, and therefore
      $\|\cdot\|_{\bm{W}}$ satisfies the 5 properties of a
      norm. However, if $\bm{W}$ were singular, the property that
      $\|\x\|_{\bm{W}} = 0 \implies \x = \bm{0}$ would be violated,
      since we could choose some $\y \neq 0 \in \vnull(\A)$ to obtain
      $\|\y\|_{\bm{W}} = \|\bm{Wy}\|_2 = \|\bm{0}\|_2 = 0$.


	\item Consider a real square matrix $\bm{M}\in\R^{n\times n}$. Suppose $\bm{M}$ is symmetric and full rank. Furthermore, suppose $\bm{M}$ is positive definite, i.e. it satisfies
	  \[
	  \forall\x\in\R^n,~\x\neq\bm{0}\implies \x^T\bm{Mx}>0.
	  \]
	  Show that the map $\la\cdot,\cdot\ra_{\bm{M}}:\R^n\to\R$ given by
	  \[
	  \la \x,\y\ra_{\bm{M}} =\x^T\bm{My}
	  \]
	  is an inner product on $\R^n$.

	  \solution Your solution.

	\item Show that the function $\la\cdot,\cdot\ra:C^0([-1,1],\R)\to\R$ given by 
	  \[
	  \la f,g\ra = \int^1_{-1}f(x)g(x)dx
	  \]
	  is an inner product on $C^0([-1,1],\R)$. Note that $C^0([-1,1],\R)$ denotes the space of continuous functions that take input from [-1,1] and produce output in $\R$.

	  \solution Your solution.

    \end{enumerate}

    \section*{Conditioning:}
    \begin{enumerate}[resume]
	\item In this problem we show that orthogonal matrices are ``perfectly conditioned'' in the sense that their condition numbers with respect to the 2-norm are always 1. Suppose $\bm{O}\in\R^{n\times n}$ is an orthogonal matrix.
	  \begin{enumerate}
	  \item Show that $\|\bm{O}\|_2=1$. (Hint: recall that orthogonal matrices preserve the 2-norms of vectors).
	  \item Show that $\bm{O}^T$ is an orthogonal matrix.
	  \item Show that $\kappa_2(\bm{O})=1$.
	  \end{enumerate}

	  \solution
	  \begin{enumerate}
	  \item Your solution.
	  \item Your solution.
	  \item Your solution.
	  \end{enumerate}
    \end{enumerate}

    \section*{Operation count:}
    \begin{enumerate}[resume]
	\item Find the number of necessary floating point operations required to compute the following operations (using the big-oh notation introduced in class). Explain your reasoning in each case.
	  \begin{enumerate}
	  \item Compute the sum $\A+\bm{B}$ for $\A,\bm{B}\in\R^{m\times n}$.
	  \item Compute the outer product $\bm{uv}^T$ for $\bm{u},\bm{v}\in\R^n$.
	  \item Compute the product $\A\x$ for $\x\in\R^n$ and $\A\in\R^{n\times n}$ where $\A$ is upper triangular.
	  \end{enumerate}

	  \solution
	  \begin{enumerate}
	  \item
        Each element of the sum requires a single addition, and there
        are $m \times n$ elements in $\A + \bm{B} \in \R^{m \times n}$, so
        this is O($m \times n$).

      \item There are  required to
        compute $\bm{uv}^T$, so this is O($n$).

      \item This is equivalent to $n$ dot-products of $n$-dimensional
        vectors. Each dot-product requires $n$ multiplies and $n-1$
        additions, or $2n - 1$ total flops. So the time complexity for
        $n$ dot-products is O($n^2$).

        Note that because $\A$ is upper-triangular, only half of the
        multiplies will be with non-zero components, but the resulting
        reduction in flops of $\half$ is irrelevant for big-oh.
	  \end{enumerate}


	\item \textbf{Comparing growth rates:} This exercise is meant to
      give you an idea of how quickly the number of flops required to
      solve a problem can increase when one increases the problem
      size, depending on the complexity of the algorithm
      used. Construct a table comparing $n,n\log_2(n),n^2,n^3,2^n,$
      and $n!$ for $n=2,4,8,16,64,512$. (You may need to use something
      like Wolfram Alpha to compute some of these quantities).

	  \solution
      \begin{tabular}{|r||r|r|r|r|r|r|}
        \hline
        & 2 & 4 & 8 & 16 & 64 & 512 \\
        \hline
        \hline
        $n$ & 2 & 4 & 8 & 16 & 64 & 512 \\
        \hline
        $n log_2(n)$ & 2 & 8 & 24 & 64 & 384 & 4608 \\
        \hline
        $n^2$ & 4 & 16 & 64 & 256 & 4096 & $2.621 \times 10^{5}$ \\
        \hline
        $n^3$ & 8 & 64 & 512 & 4096 & $2.621 \times 10^{5}$ & $1.342 \times 10^{8}$ \\
        \hline
        $2^n$ & 4 & 16 & 256 & 65536 & $1.844 \times 10^{19}$ & $1.340 \times 10^{154}$ \\
        \hline
        $n!$ & 2 & 24 & 40320 & $2.092 \times 10^{13}$ & $1.268 \times 10^{89}$ & $3.477 \times 10^{1166}$ \\
        \hline
      \end{tabular}
      
    \end{enumerate}

\end{document}
